An Improved Approximate-Bayesian Model-choice Method for Estimating Shared Evolutionary History

### 2014-02-22 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Adding divergence model plots to dpp ms.
More editing of DPP ms.

\mFigure{../../power-comparison/images/old-power-omega-4.pdf}{
    \powerCommentSummary{\powerSeriesOld}
    \powerDispersionComment{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerPsiOld4}

\mFigure{../../power-comparison/images/old-power-psi-prob-4.pdf}{
    \powerSupportCommentSummary{\powerSeriesOld}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerPsiOld4}

\mFigure{../../power-comparison/images/old-power-omega-prob-4.pdf}{
    \powerSupportCommentSummary{\powerSeriesOld}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerPsiOld4}

>>>>

\mFigure{../../power-comparison/images/old-power-omega-4.pdf}{
    \powerCommentSummary{\powerSeriesOld}
    \powerDispersionComment{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerOmegaOld4}

\mFigure{../../power-comparison/images/old-power-psi-prob-4.pdf}{
    \powerSupportCommentSummary{\powerSeriesOld}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerPsiProbOld4}

\mFigure{../../power-comparison/images/old-power-omega-prob-4.pdf}{
    \powerSupportCommentSummary{\powerSeriesOld}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerOmegaProbOld4}

%% Empirical results
\mFigure{../../empirical-analyses/plots/philippines-psi.pdf}{
    The (A--E) posterior and (F--J) prior probabilities of the number of
    divergence events (\divTimeNum) when the data of the 22 pairs of taxa from
    the Philippines are analyzed under the five models indicated at the top of
    each column of plots (Table~\ref{tabEmpiricalModels}).
    The average prior probability of an (K--O) unordered and (P--T)
    ordered model of divergence (\divTimeIndexVector) with \divTimeNum
    divergence-time parameters is also shown.
    The posterior median of the dispersion index of divergence times
    (\divTimeDispersion) is also given for each model, followed by the 95\%
    highest posterior density interval in parantheses.
}{figPhilippines}

\widthFigure{0.4}{../../empirical-analyses/plots/negros-panay-psi.pdf}{listformat=figList}{
    The posterior probabilities of the number of divergence events,
    \divTimeNum, when the data of the 9 pairs of taxa from Negros and Panay
    Islands are analyzed under the DPP model that samples over (A) unordered
    and (B) ordered models of divergence (Table~\ref{tabEmpiricalModels}).
    Both models share the same (C) prior probability of the number of
    divergence events, and the average prior probability of an (D)
    unordered and (E) ordered model of divergence (\divTimeIndexVector) with
    \divTimeNum divergence-time parameters.
    The posterior median of the dispersion index of divergence times
    (\divTimeDispersion) is also given for each model, followed by the 95\%
    highest posterior density interval in parantheses.
}{figNegrosPanay}
\clearpage

### 2014-02-22 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More editing of DPP ms.

The software package \msb implements such an approach in an
approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
The method models temporally clustered divergences across taxa caused by a
biogeographical event (or a ``divergence event'') as a single, instantaneous
occurrence.
In other words, a divergence event causes a set of taxa to share the same
moment of divergence along a continuous time scale (i.e., simultaneous
divergence).
Given aligned sequence data for \npairs{} pairs of populations, \msb 
estimates the number of divergence events shared among the pairs, the timing of
the events, and the assignment of pairs to the events, while integrating out
uncertainty in demographic parameters and the genealogical histories of the
sequences.
% samples over all the possible ways of assigning the divergences between the
% populations of each pair to 1--\npairs{} divergence events, while estimating
% the timing of the events and integrating out uncertainty in demographic
% parameters and the genealogical history of the sequences.
Thus, the method samples over all possible divergence models of differing
dimensionality (i.e., all the possible partitions of \npairs{} pairs to
$1,2,\ldots,\npairs{}$ divergence events), and, in so doing, estimates
the posterior probability of each model.

@@

\begin{linenomath}
With \alignmentVector, \hkyModelVector, \ploidyScalarVector, and
\mutationRateScalarConstantVector in hand, the joint posterior distribution
of the model is given by Bayes' rule as
\begin{equation}
    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
    \mutationRateScalarConstantVector) =
    \frac{p(\alignmentVector \given \geneTreeVector, \divTimeMapVector,
        \demographicParamVector, \locusMutationRateScalarVector,
        \locusRateHetShapeParameter, \hkyModelVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)
        p(\geneTreeVector, \divTimeMapVector, \demographicParamVector,
        \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
        \hkyModelVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)}{
        p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)},
    \label{eq:fullModelCompactJoint}
\end{equation}
which can be expanded using the chain rule of probability into components that
are assumed to be independent to get
\begin{equation}
    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
    \mutationRateScalarConstantVector) =
    \frac{p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
        \locusMutationRateScalarVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)
        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
        p(\locusRateHetShapeParameter)
        p(\divTimeMapVector)
        p(\demographicParamVector)}{
        p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)},
    \label{eq:fullModelCompact}
\end{equation}
where
$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
is a vector of population divergence times for each of the \npairs{} pairs of
populations,
$\demographicParamVector = (\demographicParams{1}, \ldots,
\demographicParams{\npairs{}})$
is a vector of the demographic parameters for each of the \npairs{} population
pairs,
$\locusMutationRateScalarVector = (\locusMutationRateScalar{1}, \ldots
\locusMutationRateScalar{\nlociTotal})$
is a vector of locus-specific mutation-rate scaling parameters for each of the
\nlociTotal loci,
\locusRateHetShapeParameter is the shape parameter of a gamma-distributed
prior on \locusMutationRateScalar{}, and
$p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
\mutationRateScalarConstantVector)$
is the probability of the data (or the marginal likelihood of the model) given
the fixed constants provided by the investigator.
\end{linenomath}

@@

\subsubsection{Likelihood terms of Equation \ref{eq:fullModelCompact}}
\begin{linenomath}
The likelihood terms of Equation \ref{eq:fullModelCompact} can
be expanded out as a product over population pairs and loci
\begin{equation}
    p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
    p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
    \locusMutationRateScalarVector, \ploidyScalarVector,
    \mutationRateScalarConstantVector) = 
    \prod_{i=1}^{\npairs{}}
    \prod_{j=1}^{\nloci{i}}
    p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})
    p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
    \locusMutationRateScalar{j}, \ploidyScalar{i}{j},
    \mutationRateScalarConstant{i}{j}).
    \label{eq:modelLikelihoodExpanded}
\end{equation}
The first term,
$p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})$,
is the probability of the sequence alignment of locus $j$ for population pair
$i$ given the gene tree and \hky substitution model parameters \citep[i.e., the
``Felsenstein likelihood'';][]{Felsenstein1981}.
The model allows for an intra-locus recombination rate \recombinationRate,
which, for simplicity, is assumed to be zero in Equation
\ref{eq:fullModelCompact}.
If \recombinationRate is non-zero, this term requires an additional product
over the columns (sites) of each sequence alignment to allow sites to have
different genealogies.
The second term,
p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
\locusMutationRateScalar{j}, \ploidyScalar{i}{j},
\mutationRateScalarConstant{i}{j}),
is the probability of the gene tree under a multi-population coalescent model
(i.e., species tree) where the ancestral population of pair $i$ diverges
and gives rise to the two sampled descendant populations.
Each \demographicParams{} contains the following demographic parameters: The
mutation-scaled effective sizes ($\myTheta{} = 4N\mutationRate$) of the
ancestral, \ancestralTheta{}, and descendant populations, \descendantTheta{1}{}
and \descendantTheta{2}{};
the proportion of the first, \bottleScalar{1}{}, and second population,
\bottleScalar{2}{}, that persist during bottlenecks that begin immediately
after divergence in forward-time;
the proportion of time between present and divergence when the bottlenecks
end for both populations, \bottleTime{};
and the symmetric migration rate between the descendant populations,
\migrationRate{}.
Thus, the probability of the $\popSampleSize{i}{}-1$ coalescence times (node
heights) of gene tree \geneTree{i}{j} is given by a multi-population
Kingman-coalescent model \citep{Kingman1982} where the ancestral population of
size
$\ancestralTheta{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
diverges at time \divTimeMap{i} into two descendant populations of constant
size
$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{1}{i}$
and
$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{2}{i}$,
which, after time $\divTimeMap{i}\bottleTime{i}$, grow exponentially to their
present size 
$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
and
$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$,
respectively.
Following divergence, the descendant populations of pair $i$ exchange migrants at a symmetric rate of
\migrationRate{i}.
\end{linenomath}

\subsubsection{Prior terms of Equation \ref{eq:fullModelCompact}}
\begin{linenomath}
The term $p(\locusRateHetShapeParameter)$ is the prior density
function for the shape-parameter of the gamma-distributed prior on
rate heterogeneity among loci.
This prior is $\locusRateHetShapeParameter \sim U(1, 20)$.
The prior probability of the vector of locus-specific mutation-rate scaling parameters
given \locusRateHetShapeParameter then expands out as a product over
the loci
\begin{equation}
    p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter) =
    \prod_{j=1}^{\nlociTotal}
    p(\locusMutationRateScalar{j} \given \locusRateHetShapeParameter),
    \label{eq:locusRateHetPrior}
\end{equation}
where each \locusMutationRateScalar{} is independently and identically
distributed (\iid) as
$\locusMutationRateScalar{} \sim Gamma(\locusRateHetShapeParameter,
1/\locusRateHetShapeParameter)$.
If the recombination rate \recombinationRate is allowed to be
non-zero, the prior term $p(\recombinationRate)$ would be added
to Equation \ref{eq:fullModelCompact}, and the prior would be
$\recombinationRate \sim Gamma(\gammaShape{\recombinationRate},
\gammaScale{\recombinationRate})$.
\end{linenomath}

>>>>

The software package \msb implements such an approach in an
approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
The method models temporally clustered divergences across taxa caused by a
biogeographical event (or a ``divergence event'') as a single, instantaneous
occurrence.
In other words, a divergence event causes a set of taxa to share the same
moment of divergence along a continuous time scale (i.e., simultaneous
divergence).
Given aligned sequence data for \npairs{} pairs of populations, \msb 
estimates the number of divergence events shared among the pairs, the timing of
the events, and the assignment of pairs to the events, while integrating out
uncertainty in demographic parameters and the genealogical histories of the
sequences.
% samples over all the possible ways of assigning the divergences between the
% populations of each pair to 1--\npairs{} divergence events, while estimating
% the timing of the events and integrating out uncertainty in demographic
% parameters and the genealogical history of the sequences.
Thus, the method samples over all possible divergence models of differing
dimensionality (i.e., all the possible partitions of \npairs{} pairs to
$1,2,\ldots,\npairs{}$ divergence-time parameters), and, in so doing, estimates
the posterior probability of each model.

@@

\begin{linenomath}
With \alignmentVector, \hkyModelVector, \ploidyScalarVector, and
\mutationRateScalarConstantVector in hand, the joint posterior distribution
of the model is given by Bayes' rule as
\begin{equation}
    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
    \mutationRateScalarConstantVector) =
    \frac{p(\alignmentVector \given \geneTreeVector, \divTimeMapVector,
        \demographicParamVector, \locusMutationRateScalarVector,
        \locusRateHetShapeParameter, \hkyModelVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)
        p(\geneTreeVector, \divTimeMapVector, \demographicParamVector,
        \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
        \hkyModelVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)}{
        p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)},
    \label{eq:fullModelCompactJoint}
\end{equation}
which can be expanded using the chain rule of probability into components that
are assumed to be independent to get
\begin{equation}
    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
    \alignmentVector, \hkyModelVector, \ploidyScalarVector,
    \mutationRateScalarConstantVector) =
    \frac{p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
        \locusMutationRateScalarVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)
        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
        p(\locusRateHetShapeParameter)
        p(\divTimeMapVector)
        p(\demographicParamVector)}{
        p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)},
    \label{eq:fullModelCompact}
\end{equation}
where
$\divTimeMapVector = (\divTimeMap{1}, \ldots, \divTimeMap{\npairs{}})$
is a vector of population divergence times for each of the \npairs{} pairs of
populations,
$\demographicParamVector = (\demographicParams{1}, \ldots,
\demographicParams{\npairs{}})$
is a vector of the demographic parameters for each of the \npairs{} population
pairs,
$\locusMutationRateScalarVector = (\locusMutationRateScalar{1}, \ldots
\locusMutationRateScalar{\nlociTotal})$
is a vector of locus-specific mutation-rate multipliers for each of the
\nlociTotal loci,
\locusRateHetShapeParameter is the shape parameter of a gamma-distributed
prior on \locusMutationRateScalar{}, and
$p(\alignmentVector \given \hkyModelVector, \ploidyScalarVector,
\mutationRateScalarConstantVector)$
is the probability of the data (or the marginal likelihood of the model) given
the fixed constants provided by the investigator.
\end{linenomath}

@@

\subsubsection{Likelihood terms of Equation \ref{eq:fullModelCompact}}
\begin{linenomath}
The likelihood terms of Equation \ref{eq:fullModelCompact} can
be expanded out as a product over population pairs and loci
\begin{equation}
    p(\alignmentVector \given \geneTreeVector, \hkyModelVector)
    p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
    \locusMutationRateScalarVector, \ploidyScalarVector,
    \mutationRateScalarConstantVector) = 
    \prod_{i=1}^{\npairs{}}
    \prod_{j=1}^{\nloci{i}}
    p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})
    p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
    \locusMutationRateScalar{j}, \ploidyScalar{i}{j},
    \mutationRateScalarConstant{i}{j}).
    \label{eq:modelLikelihoodExpanded}
\end{equation}
The first term,
$p(\alignment{i}{j} \given \geneTree{i}{j}, \hkyModel{i}{j})$,
is the probability of the sequence alignment of locus $j$ for population pair
$i$ given the gene tree and \hky substitution model parameters \citep[i.e., the
``Felsenstein likelihood'';][]{Felsenstein1981}.
The model allows for an intra-locus recombination rate \recombinationRate,
which, for simplicity, is assumed to be zero in Equation
\ref{eq:fullModelCompact}.
If \recombinationRate is non-zero, this term requires an additional product
over the columns (sites) of each sequence alignment to allow sites to have
different genealogies.
The second term,
p(\geneTree{i}{j} \given \divTimeMap{i}, \demographicParams{i},
\locusMutationRateScalar{j}, \ploidyScalar{i}{j},
\mutationRateScalarConstant{i}{j}),
is the probability of the gene tree under a multi-population coalescent model
(i.e., species tree) where the ancestral population of pair $i$ diverges
and gives rise to the two sampled descendant populations.
Each \demographicParams{} contains the following demographic parameters: The
mutation-rate-scaled effective sizes ($\myTheta{} = 4N\mutationRate$) of the
ancestral, \ancestralTheta{}, and descendant populations, \descendantTheta{1}{}
and \descendantTheta{2}{};
the proportion of the first, \bottleScalar{1}{}, and second population,
\bottleScalar{2}{}, that persist during bottlenecks that begin immediately
after divergence in forward-time;
the proportion of time between present and divergence when the bottlenecks
end for both populations, \bottleTime{};
and the symmetric migration rate between the descendant populations,
\migrationRate{}.
Thus, the probability of the $\popSampleSize{i}{}-1$ coalescence times (node
heights) of gene tree \geneTree{i}{j} is given by a multi-population
Kingman-coalescent model \citep{Kingman1982} where the ancestral population of
size
$\ancestralTheta{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
diverges at time \divTimeMap{i} into two descendant populations of constant
size
$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{1}{i}$
and
$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}\bottleScalar{2}{i}$,
which, after time $\divTimeMap{i}\bottleTime{i}$, grow exponentially to their
present size 
$\descendantTheta{1}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$
and
$\descendantTheta{2}{i}\ploidyScalar{i}{j}\mutationRateScalarConstant{i}{j}\locusMutationRateScalar{j}$,
respectively.
Following divergence, the descendant populations of pair $i$ exchange migrants at a symmetric rate of
\migrationRate{i}.
\end{linenomath}

\subsubsection{Prior terms of Equation \ref{eq:fullModelCompact}}
\begin{linenomath}
The term $p(\locusRateHetShapeParameter)$ is the prior density
function for the shape parameter of the gamma-distributed prior on
rate heterogeneity among loci.
This prior is $\locusRateHetShapeParameter \sim U(1, 20)$.
The prior probability of the vector of locus-specific mutation-rate multipliers
given \locusRateHetShapeParameter then expands out as a product over
the loci
\begin{equation}
    p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter) =
    \prod_{j=1}^{\nlociTotal}
    p(\locusMutationRateScalar{j} \given \locusRateHetShapeParameter),
    \label{eq:locusRateHetPrior}
\end{equation}
where each \locusMutationRateScalar{} is independently and identically
distributed (\iid) as
$\locusMutationRateScalar{} \sim Gamma(\locusRateHetShapeParameter,
1/\locusRateHetShapeParameter)$.
If the recombination rate \recombinationRate is allowed to be
non-zero, the prior term $p(\recombinationRate)$ would be added
to Equation \ref{eq:fullModelCompact}, and the prior would be
$\recombinationRate \sim Gamma(\gammaShape{\recombinationRate},
\gammaScale{\recombinationRate})$.
\end{linenomath}

### 2014-02-22 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
Working on dpp manuscript.
Proofreading and incorporating edits throughout the manuscript.
Changing first person to singular in dpp ms.
More editing of DPP ms.

        \divTimeNum & Number of unique population divergence times among the \npairs{} population pairs. \tn
        \divTime{} & One of the \divTimeNum \iid divergence time paramters in \globalcoalunit generations. \tn
        \divTimeVector & Ordered set of divergence time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
        \divTimeIndex{i} & The index of the divergence time parameter in \divTimeVector that is mapped to population pair $i$. \tn
        \divTimeIndexVector & Vector of divergence time parameter indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn

@@

        \descendantTheta{1}{i}, \descendantTheta{2}{i} & \myTheta{} of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
        \ancestralTheta{i} & \myTheta{} of the population ancestral to pair $i$. \tn

@@

        \locusMutationRateScalar{j} & mutation-rate scaling parameter for locus $j$. \tn
        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate parameters: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
        \locusRateHetShapeParameter & Hyper-parameter for the shape of the gamma prior distribution on each \locusMutationRateScalar{}. \tn

@@

        \bottleTime{i} & Proportion of time between present and \divt{i} when the bottleneck ends for the descendant populations of pair $i$. \tn

@@

        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$. This constant is intended to allow the user to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa. \tn
        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$. This constant is intended to allow the user to scale \myTheta{} for differences in mutation rates among loci or among taxa. \tn

@@

        \divTimeMean  & Mean of \divTimeMapVector. \tn
        \divTimeVar  & Variance of \divTimeMapVector. \tn
        \divTimeDispersion  & Dispersion index of \divTimeMapVector ($\divTimeVar/\divTimeMean$). \tn

@@

\begin{table}[htbp]
    \sffamily
    \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The model priors evaluated in our simulation-based analyses.  For
        the \modelDPP model, the prior on the concentration parameter,
        $\concentrationParam \sim Gamma(\cdot,\cdot)$, was set to $Gamma(2,2)$
        for the validation analyses and $Gamma(1.5,18.1)$ for the power
        analyses. The distributions of divergence times are given in units of
        \globalcoalunit generations followed in brackets by units of millions
        of generations ago (MGA), with the former converted to the latter
        assuming a per-site rate of 1\e{-8} mutations per generation. For model
        \modelOld, the priors for theta parameters are $\ancestralTheta{} \sim
        U(0, 0.05)$ and $\descendantTheta{1}{}, \descendantTheta{2}{} \sim
        Beta(1, 1) \times 2 \times U(0, 0.05)$. The later is summarized as
        $\descendantThetaMean{} \sim U(0, 0.05)$. For the \modelDPP and
        \modelUniform, and \modelUshaped models, \ancestralTheta{},
        \descendantTheta{1}{}, and \descendantTheta{2}{} are independently
        and exponentially distributed with a mean of 0.025.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & \multicolumn{4}{c}{Priors} \\
        \cmidrule(){2-5}
        Model & \divTimeIndexVector & \divTime{} & \myTheta{} &  \\
        \midrule
            \modelOld & $\divTimeIndexVector \sim \priorOld$
                      & $\divTime{} \sim U(0,10 \; [25 \; MGA])$
                      & $\ancestralTheta{} \sim U(0, 0.05)$
                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
                      & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                      & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
                          & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                          & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
                          & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                          & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
        \bottomrule
    \end{tabular}
    \label{tabPriors}
\end{table}

@@

    \caption{The models we used to simulate pseudo-replicate datasets for

>>>>

        \divTimeNum & Number of population divergence-time parameters shared among the \npairs{} population pairs. \tn
        \divTime{} & Time of population divergence in \globalcoalunit generations. \tn
        \divTimeVector & Set of divergence-time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
        \divTimeIndex{i} & The index of the divergence-time in \divTimeVector to which population pair $i$ is mapped. \tn
        \divTimeIndexVector & Vector of divergence-time indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn

@@

        \descendantTheta{1}{i}, \descendantTheta{2}{i} & Mutation-rate-scaled effective population size of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
        \ancestralTheta{i} & Mutation-rate-scaled effective population size of the population ancestral to pair $i$. \tn

@@

        \locusMutationRateScalar{j} & Mutation-rate multiplier of locus $j$. \tn
        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate multipliers: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
        \locusRateHetShapeParameter & The shape parameter of the gamma prior distribution on \locusMutationRateScalar{}. \tn

@@

        \bottleTime{i} & Proportion of time between present and \divTimeMap{i} when the bottleneck ends for the descendant populations of pair $i$. \tn

@@

        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the investigator for locus $j$ of pair $i$. This constant is required to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa. \tn
        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the investigator for locus $j$ of pair $i$. This constant is required to scale \myTheta{} for differences in mutation rates among loci or among taxa. \tn

@@

        \divTimeMean  & Mean of divergence times across the \npairs{} population pairs. \tn
        \divTimeVar  & Variance of divergence times across the \npairs{} population pairs. \tn
        \divTimeDispersion  & Dispersion index of divergence times across the \npairs{} population pairs ($\divTimeVar/\divTimeMean$). \tn

@@

\begin{table}[htbp]
    \sffamily
    \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The models evaluated in the simulation-based analyses.  For
        the \modelDPP model, the prior on the concentration parameter,
        $\concentrationParam \sim Gamma(\cdot,\cdot)$, was set to $Gamma(2,2)$
        for the validation analyses and $Gamma(1.5,18.1)$ for the power
        analyses. The distributions of divergence times are given in units of
        \globalcoalunit generations followed in brackets by units of millions
        of generations ago (MGA), with the former converted to the latter
        assuming a per-site rate of 1\e{-8} mutations per generation. For model
        \modelOld, the priors for theta parameters are $\ancestralTheta{} \sim
        U(0, 0.05)$ and $\descendantTheta{1}{}, \descendantTheta{2}{} \sim
        Beta(1, 1) \times 2 \times U(0, 0.05)$. The later is summarized as
        $\descendantThetaMean{} \sim U(0, 0.05)$. For the \modelDPP and
        \modelUniform, and \modelUshaped models, \ancestralTheta{},
        \descendantTheta{1}{}, and \descendantTheta{2}{} are independently
        and exponentially distributed with a mean of 0.025.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & \multicolumn{4}{c}{Priors} \\
        \cmidrule(){2-5}
        Model & \divTimeIndexVector & \divTime{} & \myTheta{} &  \\
        \midrule
            \modelOld & $\divTimeIndexVector \sim \priorOld$
                      & $\divTime{} \sim U(0,10 \; [25 \; MGA])$
                      & $\ancestralTheta{} \sim U(0, 0.05)$
                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
                          & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                          & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
                          & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                          & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
                      & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                      & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
        \bottomrule
    \end{tabular}
    \label{tabPriors}
\end{table}

@@

    \caption{The models used to simulate pseudo-replicate datasets for
